### 2015-12-31 netft.tex
Initial commit.

\documentclass[twocolumn]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{breqn}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

\title{DRAFT: Robust Communication in Structured Networks}
\author{
    Edward L. Platt\\
    elplatt@umich.edu
    \and
    Daniel M. Romero\\
    drom@umich.edu
}
\date{December 2015}

\begin{document}

\maketitle

\section{Introduction}

\section{Trust Model}

\section{Fault Tolerance}

\section{Butterfly Network}

\section{Discussion}

\section{Conclusion}

\bibliographystyle{plain}
\bibliography{netft}

\end{document}