### 2016-02-18 ec-2016/source/paper.tex
Add partial draft of EC16 writeup.

%\documentclass[prodmode]{acmsmall-ec16}
%\documentclass[prodmode,license]{acmsmall-ec16}
\documentclass[prodmode,permissions]{acmsmall-ec16}

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\usepackage{amsmath,amsfonts,amssymb,bbm} 
\usepackage[numbers,sort&compress]{natbib} % for citet
\SetArgSty{textrm}  % for algorithm2e
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%\conferenceinfo{EC'16,}{July 24--28, 2016, Maastricht, The Netherlands.}
%\CopyrightYear{2016}
%\crdata{978-1-4503-3936-0/16/07}
%\copyrighttext{Copyright is held by the owner/author(s). Publication rights licensed to ACM.}

\doi{XXXXXXX.XXXXXXX}% TeXSupport

\newcommand{\beq}{\begin{eqnarray}}
\newcommand{\eeq}{\end{eqnarray}}
\DeclareMathOperator{\wbf}{wBF}

% Document starts
\begin{document}

% Page heads
%\markboth{G. Zhou et al.}{A Multifrequency MAC Specially Designed for WSN Applications}

% Title portion
\title{
Attack-Tolerance in Structured Networks via Multipath Routing
\\ Draft: 2016-02-16
}
\author{EDWARD L. PLATT
\affil{University of Michigan}
DANIEL M. ROMERO
\affil{University of Michigan}}

\begin{abstract}
TODO
\end{abstract}

%ACM is moving forward with the 2012 Classification system: http://dl.acm.org/ccs_flat.cfm. Please generate the CCSXML tex code through the online interactive system and insert the code below.

\begin{CCSXML}
<ccs2012>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{TODO~TODO}

%\terms{Design, Algorithms, Performance}

\keywords{
censorship,
decentralization,
fault tolerance,
multipath,
networks,
peer-to-peer,
trust
}

\begin{bottomstuff}
This work is supported by TODO.

Author’s addresses: E.L. Platt and D.M. Romero, School of Information; email: \{elplatt,drom\}@umich.edu.
\end{bottomstuff}

\maketitle

TODO: Fix bibtex encoding.

\section{Introduction}

Communication networks, expemplified by the Internet,
have become ubiquitous and critical infrastructural for
communities, organizations, and markets.
As with any critical infrastructure, the cost of a failure can be
immense, so methods for tolerating various kind of faults are an
important and ongoing area of research.
The key question is: what are the techniques and network structures that can
be used to create robust, fault-tolerant systems?

Many complex networks, including the Internet and World-Wide Web, exhibit
scale-free structure
\cite{barabasi_emergence_1999,barabasi_scale-free_2009}.
While scale-free networks tolerate random faults well,
they are highly susceptible to adversarial faults, i.e. attacks
\cite{albert_error_2000}.
For example, in 2008, YouTube suffered a worldwide outage for several hours
when a service provider in Pakistan advertised false routing information
\cite{hunter_pakistan_2008}.
The action (known as a {\em black hole attack}) was intended to censor YouTube
within Pakistan only, but resulted in a worldwide cascading failure.
Such vulnerabilities are not limited to any one system or protocol, but an
attribute of complex communication networks themselves.
General techniques for understanding and mitigating these vulnerabilities are
needed.

In this paper, we formally evaluate the effects of network structure on
attack-resistance. We also propose an efficient, fault-tolerant network architecture.
Our analysis focuses on {\em structured networks},
in which links between nodes are constrained to a particular architecture.
The proposed fault-tolerant architecture utilizes {\em multipath routing},
in which many possible paths exist betweeen two nodes on a communication network.
In our case, multiple paths allow multiple copies of a message to be sent,
which can be used for error-detection and error-correction by the receiver.

Our main contributions are:
\begin{itemize}
\item{We present a formal {\em partial trust model},
which makes weaker transitivity assumptions than previous web-of-trust models,
and use this model to quantify the effects of network structure on
fault tolerance;}
\item{We show that the probability of successfully detecting an
adversarial fault increases exponentially with the number of disjoint untrusted
paths between the neighborhoods of the sender and receiver, and that this
value depends on network structure;}
\item{We present a scalable, efficient, and attack-tolerant multipath
routing algorithm on the butterfly network topology.}
\end{itemize}

Paper organization. TODO

\section{Background and Related Work}

In {\em structured networks}, link structure is predetermined.
Such networks can be designed to have favorable structural and routing properties,
at the expense of complicating the addition or removal of nodes.
Structured networks have been a popular tool in parallel and distributed
computing architectures \cite{kshemkalyani_distributed_2008}.
More recently, peer-to-peer systems based on distributed hash tables have used
structured ``overlay'' networks to map table keys to local TCP/IP routes
\cite{lua_survey_2005}.

Many distributed consensus protocols (such as those used by crypto-currencies)
are designed to provide tolerance against arbitrary or adversarial faults.
Byzantine agreement protocols
\cite{lamport_byzantine_1982,castro_practical_1999}
provide tolerance against arbitrary faults (including attacks) under
some circumstances, but are limited to small networks due to poor scalability.
Proof-of-work \cite{dwork_pricing_1993,nakamoto_bitcoin:_2008}
and proof-of-stake \cite{king_ppcoin:_2012}
provide better scalability, but wasteful of computational and energy
resources.
Federated Byzantine Agreement (FBA) \cite{mazieres_stellar_2015}
is scalable and optimal, guaranteeing that the protocol only fails when
success is impossible, although it does not provide a way to evaluate
the probability of failure.
All of these protocols rely on the assumption that their cryptography
cannot be compromised.
While cryptography can be extremely resistant to technological attacks,
it can still be compromised through coercion (e.g., legal action).
In additon, the fault-tolerance of FBA depends on redundancy in the network structure
but leaves an unanswerd question: how should a network be structured to achieve
fault tolerance?
This paper addresses both of these issues by analyzing the role of network
structure on adversarial faults, without relying on assumptions of cryptographic
integrity.

{\em Multipath} routing uses multiple paths when routing a message through a
network, in contrsast to traditional {\em unipath} routing, which uses
a single path.
Multipath routing can have many benefits, including reduced congestion,
increased throughput, and more reliability
\cite{qadir_exploiting_2015}.
Many of these routing protocols offer better security 
\cite{zin_survey_2015}.
Some approaches utilize redundant paths as backups for increased
fault tolerance
\cite{alrajeh_secure_2013},
and some specificially protect against adversarial faults
\cite{kohno_improvement_2012, khalil_unmask:_2010, lou_h-spread:_2006}.
The method of Liu et al.
\cite{liu_secure_2012}
routes multiple messages first to random peers and then
to their final destination.
The butterfly algorithm we present takes a conceptually similar approach.
Most work on multipath routing has been motivated by applications related to
wireless sensor networks (WSNs),
and have thus focused on ad-hoc, unstructured networks, often having a central
base station.
The trust model presented in this paper provides a more general way to evaluate
the effect of network structure on adversarial fault-tolerance.

\section{Attack-Tolerant Networks Infrastructure}

In this section, we describe the functional proprties required of an
attack-tolerant network infrastructure and how those properties translate
into constraints on network structure.
We pay special attention to a property we call
{\em stabilizing asymmetry}.

\subsection{Functional Properties}

In most cases, infrastructure must remain functional as it grows;
it must be {\em scalable}.
Systems having single points of failure are less tolerant against faults at those
points, which both raises the likelihood of a failure and creates targets
where attackers might be able to focus their resources efficiently.
Attack-tolerant infrastructure must minimize single points of failure;
it must be {\em decentralized}.
In some, but not all, cases it is also desirable that an infrastructure
provides {\em secrecy},
protecting the contents of a message or the identity of the sender.

We add one additional property, which we call {\em stabilizing asymmetry}.
In the context of international conflict,
\cite{mack_why_1975}
observed that power imbalance usually determines the outcome of a conflict
(with the more powerful side winning)
except in the special case of {\em asymmetric conflicts}.
In asymmetric conflicts, the same level of resource expenditure yields different
results for different parties.
There are two possible cases:
the attacker's resources are either more or less effective than the defender's.
We call the latter case stabilizing asymmetry,
because it lowers the incentive to attack.
With this in mind, an attack-resistant infrastructure will benefit from a high
level of stabilizing asymmetry.

\subsection{Structural Properties}

Sparse

Low-diameter

Vertex-transitivity

Redundancy
\cite{baran_distributed_1964}

Routing

\subsection{Trust}

Scale and indirect communication

Source-to-network
\begin{figure}
\centerline{\includegraphics{fig-trust-source}}
\caption{TODO}
\label{fig:trust-source}
\end{figure}

Source-to-destination
\begin{figure}
\centerline{\includegraphics{fig-trust-source-destination}}
\caption{TODO}
\label{fig:trust-source-destionation}
\end{figure}

\section{Multipath Fault Tolerance}

Fault tolerance with partial trust and redundancy.

\subsection{Fault Tolerance}

Results

\begin{figure}
\centerline{\includegraphics{fig-perror}}
\caption{TODO}
\label{fig:one}
\end{figure}

\section{Multipath Routing on the Butterfly Topology}

We now describe a Concurrent Multipath Routing scheme on a specific family
of networks: the butterfly network \cite{}.
Several variations on the butterfly network exist.
Specifically, we utilize the wrap-around butterfly.
We denote the $m$-dimensional directed wrap-around butterfly as $\wbf(m)$:
\beq
\wbf(m) &=& (V, E_\downarrow \cup E_\rightarrow) \\
V &=& \mathbb{Z}_m \times \mathbb{Z}_{2^m} \\
E_\downarrow &=& \{((l,z),(l+1 (\text{mod } m),z) \nonumber \\
&& \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\} \\
E_\rightarrow &=& \{(l,z),(l+1 (\text{mod } m), z \oplus 2^l) \nonumber \\
&& \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\},
\eeq
where $\oplus$ represents the bitwise XOR operator.
Each node is associated with a level $l$ and an $m$-bit integer $z$.
There are two types of edges, shown in Figure \ref{fig:butterfly}.
Down edges ($E_\downarrow$) connect nodes sharing the same $z$ value
in a cycle of increasing level $l$.
Down-right edges ($E_\rightarrow$) also link to a node of level $l + 1$,
but one having the bitstring equal to $z$ with the $l$th bit flipped.

The wrap-around butterfly network has a number of properties making it useful for multipath routing:
\begin{description}
\item[Vertex transitivity]
The problem
of finding a route between arbitrary nodes $v$ and $w$
can be reduced to finding a route from node $(0,0)$ to some $\tilde{w}$.
\item[Logarithmic diameter]
For any two nodes, the length of the shortest path between them is logorithmic
in the size of the network.
\item[Constant degree]
In practical applications, each communication link requires additional resources,
such as physical infrastructure or entries in a routing table.
A constant degree limits the number of such resources needed as the network
grows in size.
\end{description}

\begin{figure}
\begin{center}
\includegraphics{fig-butterfly.pdf}
\end{center}
\caption{
Schematic illustration of the two types of edges in a directed butterfly network.
The node $(l,z)$ is shown as the bit string $z$ with a square around the $l$th bit.
\label{fig:butterfly}
}
\end{figure}

We now describe a routing scheme that provides $2^h$ redundant paths between
the $h$-hop trusted neighborhoods of any two nodes in an $m$-bit
wrap-around butterfly network.
Utilizing vertex transitivity, we label the source node as $(0, 0)$ and
denote the destination node as $w = (l_w, z_w)$.

Let $s$ be an integer such that $0 \leq s < 2^h$.
Let $v_s^{(t)} = (l^{(t)}, z^{(t)})$ be the $t$th node in the path labeled by $s$.
For convenience, we will omit the subscript $s$.
We define two partitionings of the integers in $\mathbb{Z}_{2^m}$:
one having the lowest $h$ bits matching the
bits of $s$, and one having the $h$ bits preceeding the destination level $l_w$
matching $s$:
\beq
S_s &=& \{z \in \mathbb{Z}_{2^m} | \forall i \in \mathbb{Z}_h z_i = s_i \} \\
R_s &=& \{z \in \mathbb{Z}_{2^m} | \forall i \in \mathbb{Z}_h z_{(l_w - h + i)} = s_i \}.
\eeq
Note that if $r \neq s$, then $S_s \cap S_r = R_s \cap R_r = \emptyset$.

We now construct a path such that between trusted neighborhoods $z^{(t)}$ is always
in $S_s$, $R_s$, or both, guaranteeing that the path does not overlap with the
other paths $v_r^{(t)}$.
Routing proceeds in stages, with the level $l$ increasing by 1 at each hop.
In Stage 1 ($0 \leq t < h$), down or down-right edges
are chosen such that the the $t$th bit of $z^{(t+1)}$ is equal to the $t$th bit
of $s$. Throughout Stage 1, all nodes are within the sender's trusted neighborhood.
At the end of Stage 1, $z^{(h)} \in S_s$, and $z^{(t)}$ will remain so until the level loops
back to $0$ at $t = m$.

In Stage 2 ($h \leq t < l_w - h$), edges are chosen to make the $t$th bit of
$z^{(t+1)}$ match the $t$th bit of $z_w$.

In Stage 3 ($l_w - h \leq t < l_w$), the bits of $z^{(t)}$ are chosen to match $s$,
such that after the stage is complete, $z^{(l_w)} \in R_s$.

In Stage 4 ($l_w \leq t < m$), as in stage 2,
paths are chosen such that the $t$th bit of $z^{(t+1)}$ matches $z_w$.
After Stage 4, all bits of $z^{(m)}$ are equal to those of $z_w$
except for the first $h$ and the $h$ preceeding index $l_w$.
$z^{(m)}$ is also in both $S_s$ and $R_s$.

At this point, we define $\tau = t - m$.
In Stage 5 ($0 \leq \tau < h$), the first $h$ bits of $z^{(t)}$ are set to
match $z_w$, potentially removing $z^{(t)}$ from $S_s$.

In Stage 6 ($h \leq \tau < l_w - h$), all down edges are chosen, incrementing
the level without any effect on $z^{(t)}$.
At the end of Stage 6, $z^{(m + l_w - h)}$ is still in $R_s$ and
$v^{(m + l_w - h)}$ is now within the trusted neighborhood of $w$.

In the seventh, and final stage ($l_w - h \leq \tau < l_w$), the $h$ bits of $z^{(t)}$
preceeding index $l_w$ are set to match $z_w$.
After this stage, $v^{(m + l_w)} = w$ and routing is complete.

\begin{figure}
\begin{center}
\includegraphics{fig-routing.pdf}
\end{center}
\caption{
\label{fig:routing}
}
\end{figure}

\section{Discussion}

Secrecy

Creation of the network.
Improves on web of trust.
Gives framework for determining where to build trust.

Applications
Distributed apps: storage, email, cryptocurrency, secure multiparty computation.
Wireless Sensor Networks

\section{Conclusion}

% Acknowledgments
\begin{acks}
The authors would like to thank TODO
\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{paper}

\end{document}