### 2021-09-19 NetDelibSim/main.tex
Move main text of ABM paper into proposal.

\documentclass[twocolumn,10pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{times}
\usepackage{geometry}

\usepackage[switch]{lineno}
\linenumbers

\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{claim}{Claim}

\DeclareMathOperator{\unif}{unif}
\DeclareMathOperator{\sample}{sample}
\DeclareMathOperator{\setcount}{count}
\DeclareMathOperator{\mode}{mode}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\usepackage{geometry}
\geometry{left=1.5cm,right=1.5cm,top=2cm,bottom=2cm}
\setlength{\columnsep}{0.5cm}

\title{Small interlocking groups improve mass deliberation\\in the presence of strong social influence}
\author{
%Edward L. Platt\\
%University of Michigan\\
%elplatt@umich.edu
%and
%Herminio Bodon\\
%Northwestern University\\
%HerminioBodon2020\\@u.northwestern.edu
%\and
%Daniel M. Romero\\
%University of Michigan\\
%drom@umich.edu
Authors Redacted \\
for \\
Double-Blind Peer-Review
}
\date{\today}

\begin{document}

\maketitle

\bibliography{references}
\bibliographystyle{plain}

\end{document}