### 2021-09-19 acknowledgements.tex

Thanks to Daniel M. Romero.

Scott E. Page,
Tawanna Dillahunt,
Ceren Budak,
for their time, guidance, and helpful feedback.

Jane Im,

Herminio Bodon,

Danielle Livneh,
Karthik Ramanathan
for help collecting WikiProjects data.

Yan Chen,
Tanya Rosenblat,
Patrick Grim,

Grant Schoenbeck,
Kentaro Toyama,
Charles R. Severance,

Veronica Falandino,
Linda Wood,
Mita Gibson,
Denise Bailey,
Liz Ratzloff,
Salem Elzway,

Jeffrey W. Lockhart,
Danaja Maldeniya,
Lia Bozarth,
Christopher Quarles,
Chanda Phelan,
Padma Chirumamilla,
Earnest Wheeler,
Sam Carton,
Elliott Brannon,
Tawfiq Ammari,
2015 UMSI PhD chohort
Jean Hardy,
Cindy Lin,

the attendees of the May 25, 2017 MIT Center for Civic Media lab meeting
the Berkman-Klein Center Cooperation Working Group
the 2019 CSCW Workshop on Team and Group Diversity
for helpful feedback;

This research was partly supported by
the National Science Foundation under Grant No. IIS-1617820,
Air Force Office of Scientific Research under Award No. FA9550-19-1-0029,
the UMSI REMS program,

Andy Brosius,
Persephone Hernandez-Vogt,

Anthony Garnock-Jones,
Elaine Lee,

i3 Detroit,

Rahul Bhargava,
Ethan Zuckerman,
Sasha Costanza-Chock,
Willow Idlewild,
Charlie DeTar,

Gerald J. Sussman