### 2021-09-19 proposal.tex

\documentclass{report}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{apacite}
\usepackage{bbding}
\usepackage{courier}
\usepackage{epigraph}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage[utf8]{inputenc}
\usepackage[switch]{lineno}
\usepackage[normal]{threeparttable}
\usepackage{titlesec}
\usepackage{url}

\newcommand{\beq}{\begin{eqnarray}}
\newcommand{\eeq}{\end{eqnarray}}
\newcommand{\+}{\phantom{-}}

\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{claim}{Claim}

\DeclareMathOperator{\unif}{unif}
\DeclareMathOperator{\sample}{sample}
\DeclareMathOperator{\setcount}{count}
\DeclareMathOperator{\mode}{mode}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\linenumbers

\author{Edward L. Platt}
\title{Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making}

\begin{document}

\maketitle

\begin{abstract}
\include{abstract.tex}
\end{abstract}

\tableofcontents

\chapter{Introduction}
\include{ch-intro.tex}

\chapter{Network structure, productivity, and performance in WikiProjects}
\include{ch-wp-prod-perf.tex}

\chapter{Small interlocking groups improve mass deliberation in the presence
of strong social influence}
\include{ch-abm.tex}

\chapter{Experimental evaluation of Network Deliberaiton}

\chapter*{Acknowledgements}
%\include{acknoledgements.tex}

\nocite{*}
\bibliographystyle{apacite}
\bibliography{proposal}

\end{document}