### 2021-09-28 proposal.tex
Figures and title page.

\documentclass{report}

\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{apacite}
\usepackage[toc,page]{appendix}
\usepackage{bbding}
\usepackage{courier}
\usepackage{epigraph}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage[utf8]{inputenc}
\usepackage[switch]{lineno}
\usepackage[normal]{threeparttable}
\usepackage{titlesec}
\usepackage{url}

\newcommand{\beq}{\begin{eqnarray}}
\newcommand{\eeq}{\end{eqnarray}}
\newcommand{\+}{\phantom{-}}

\renewcommand{\labelenumii}{\theenumii}
\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}

\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{claim}{Claim}

\DeclareMathOperator{\unif}{unif}
\DeclareMathOperator{\sample}{sample}
\DeclareMathOperator{\setcount}{count}
\DeclareMathOperator{\mode}{mode}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

%\linenumbers

\author{Edward L. Platt}
\title{Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making}

\begin{document}

\begin{titlepage}
   \begin{center}
       \vspace*{1cm}

       \textbf{\Large Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making}

       \vspace{1.5cm}

       \textbf{Edward L. Platt}

       \today

       \vfill

       Dissertation Proposal\\
       Doctor of Philosophy \\
       University of Michigan \\
       School of Information

       \vspace{0.8cm}

   \end{center}
    {\bfseries Doctoral Committee:}\\
    Associate Professor Daniel M. Romero \\
    Professor Scott E. Page \\
    Associate Professor Tawanna Dillahunt \\
    Assistant Professor Ceren Budak

\end{titlepage}

%\begin{abstract}
%\include{abstract.tex}
%\end{abstract}

\tableofcontents

\listoffigures
\listoftables

\chapter{Introduction}
\label{chap:intro}
%\include{ch-intro.tex}

\chapter{Network structure, productivity, and performance in WikiProjects}
\label{chap:wp-prod-perf}
%\include{ch-wp-prod-perf.tex}

\chapter{Small interlocking groups improve mass deliberation in the presence
of strong social influence}
\label{chap:abm}
%\include{ch-abm.tex}

\chapter{Experimental evaluation of Network Deliberaiton}
\label{chap:experiment}
\include{ch-experiment.tex}

%\chapter*{Acknowledgements}
%\include{acknoledgements.tex}

\begin{appendices}

\chapter{The NK-Model}
\label{apx:nk}
\include{apx-nk.tex}

\chapter{Interview Script}
\label{apx:interview}
\include{apx-interview.tex}

\end{appendices}

\nocite{*}
\bibliographystyle{apacite}
\bibliography{proposal}

\end{document}