### 2022-05-02 Proposal/apx-interview.tex

\section{Preamble}

Hi! Thank you so much for taking time out to talk to us today. We appreciate it. Today we are here to understand how you use the internet to discuss complex issues and your experience participating in our discussion study.

\section{Warm-Up}

\begin{enumerate}
\item How often do you discuss political and policy issues with others, in general?
\item Where and how do you discuss those political and policy issues?
\item Which online social networks, forums, email lists, etc. do you regularly participate in?
\item In general, what have your experiences discussing politics and policy online been like?
\item Do you have any memorable experiences discussing an issue online?
\end{enumerate}

\section{Main Questions}

\begin{enumerate}
\item Reminder of topic, number of stages, length of stages
\item Going into the deliberation, what was your goal?
\item Overall, how would you describe your experience?
\item Do any memorable moments stand out to you?
\item How did it compare to other group deliberations you have participated in? 
\item What was your opinion in at the beginning of the deliberation?
\item What types of opinions did you encounter from other participants?
\begin{enumerate}
\item (follow-up) Can you walk me through how your opinions evolved throughout the deliberation?
\end{enumerate}
\item How much consensus existed at the beginning of the deliberation?
\begin{enumerate}
\item (follow-up) Can you walk me through how the group consensus evolved throughout the deliberation? 
\end{enumerate}
\item Did you witness or experience disagreement throughout the study?
\begin{enumerate}
\item (follow-up) Can you describe how disagreements and conflicts were addressed during deliberation? 
\item (follow-up) What is your opinion of the way disagreements and conflicts were addressed? (If clarification needed: was the process pleasant or unpleasant? Was it effective?)
\end{enumerate}
\item Did you share all of your thoughts with your groups?
\begin{enumerate}
\item (If the participant answers no) What do you think were the reasons?
\end{enumerate}
\item How often did you read posts by other participants? 
\item How often did you read replies to posts by other participants?
\item How often did you reply to posts by other participants?
\item What was your experience like when replying to previous comments?
\end{enumerate}

\section{Debrief}

\section{Final Questions}

\begin{enumerate}
\item Do you have any feedback on how to improve the system?
\end{enumerate}