### 2022-07-05 UMich/front_materials/acknowledgements.tex
Update acknowledgements and dedication.

Thanks to my advisor, Daniel M. Romero for his time, expertise, guidance, and patience.
Thanks to my committee:
Scott E. Page,
Tawanna Dillahunt,
Ceren Budak,
for their time, expertise, and feedback.

Thanks to Danielle Livneh, and Karthik Ramanathan for help collecting the WikiProjects data used in Chapter \ref{chap:wp-prod-perf}.
Thanks to Hermino Bodon for collaboration on the agent-based modeling code in Chapter \ref{chap:abm}.
Thanks to Jane Im for collaboration on interviewing the participants in the experiment described in Chapter \ref{chap:experiment}.
Thanks to Nicole Ellison, Yan Chen, Tanya Rosenblat, and Patrick Grim for feedback on projects and proposals that led to the work in this dissertation.
Thanks to Grant Schoenbeck, Kentaro Toyama, and Charles R. Severance for feedback and insightful discussion.

Chapter \ref{chap:wp-prod-perf} has been adapted from previously published work \cite{platt_network_2018}.

This research was partly supported by
the National Science Foundation under Grant No. IIS-1617820,
Air Force Office of Scientific Research under Award No. FA9550-19-1-0029,
and the UMSI REMS program.

Thanks to Veronica Falandino for making UMSI a kinder place.
Thanks to Linda Wood and Mita Gibson of the Center for the Study of Complex Systems always being willing to help. 
Thanks to the members and staff of the Graduate Employees Organization, AFT Michigan 3550, especially Denise Bailey and Liz Ratzloff, for advocating for the well-being of Michigan graduate students.
Thanks to Avdi Grimm for consulting on some very challenging Ruby bugs.
Thanks to Salem Elzway for technical writing feedback.

Thanks to many Michigan graduate students for feedback and insightful conversations, including:
Jeffrey W. Lockhart,
Danaja Maldeniya,
Lia Bozarth,
Christopher Quarles,
Chanda Phelan,
Padma Chirumamilla,
Earnest Wheeler,
Sam Carton,
Elliott Brannon,
Tawfiq Ammari,
and the 2015 UMSI PhD chohort,
in particular Jean Hardy and Cindy Lin.
Thanks to attendees of the May 25, 2017 MIT Center for Civic Media lab meeting,
the Berkman-Klein Center Cooperation Working Group,
and the 2019 CSCW Workshop on Team and Group Diversity
for helpful feedback.
Thanks to
Lorrie LeJeune,
Rahul Bhargava,
Ethan Zuckerman,
Sasha Costanza-Chock,
Willow Idlewild,
Charlie DeTar,
Erhardt Graeff,
J. Nathan Matias,
Gerald J. Sussman,
Anthony Garnock-Jones,
Brandon Willard,
Laurel Ruhlen,
and Elaine Lee
for guidance and encouragement in returning to academia.
Thanks to the past and present members of i3 Detroit for inspiring my interest in participatory governance.

Thanks to Caitlin Cassady and Justin Veenhuis for taking in strays.
Thanks to Persephone Hernandez-Vogt for their companionship and encouragement.
Thanks to Andy Brosius for sharing their life and family with me, and for their many years of patience and encouragement.